other functionalities

### 2024-11-12 inst/gitbook/03_specific-tutorials.Rmd
book up

One other argument is `dimensions` that specifies how many dimensions from the matrix should be used for further analysis.
When missing, all dimensions from the ordinated matrix are used.

```{r, eval=TRUE}
## Using the first 50% of the dimensions
boot.matrix(BeckLee_mat50, dimensions = 0.5)

## Using the first 10 dimensions
boot.matrix(BeckLee_mat50, dimensions = 10)
```
Of course, one could directly supply the subsets generated above (using `chrono.subsets` or `custom.subsets`) to this function.

@@

```{r, fig.width=8, fig.height=4}
## Making the different subsets
cust_subsets <- custom.subsets(BeckLee_mat99,
                               crown.stem(BeckLee_tree,
                                          inc.nodes = TRUE))
time_subsets <- chrono.subsets(BeckLee_mat99,
                               tree = BeckLee_tree,
                               method = "discrete",
                               time = 5)

@@

```{r, fig.height = 16, fig.width = 8}
par(mfrow = c(2,1))
## Default plotting
plot(disparity_time_slices, main = "Disparity through time")
## Plotting with more arguments
plot(disparity_time_slices, type = "preview",
     main = "Two first dimensions of the trait space")
```

@@

@@

```{r, fig.width=8, fig.height=8}
## Graphical parameters
op <- par(bty = "n", mfrow = c(1, 2))

@@

```{r, print = FALSE}
par(mfrow = c(3,1))
slices <- c(7.9, 3.95, 0)
fun.plot <- function(tree) {
  plot(tree)
  nodelabels(tree$node.label, cex = 0.8)
  axisPhylo()
  abline(v = tree$root.time - slices)
}
silent <- lapply(trees, fun.plot)
```

@@

Variance-covariance matrices are sometimes a useful way to summarise multidimensional data.
In fact, you can express the variation in your multidimensional dataset directly in terms of how your trait covary rather than simply the positions of your elements in the trait space.
Furthermore, variance-covariance matrices can be estimated from multidimensional in sometimes more useful ways that simply looking at the the data in your trait space.
This can be done by describing your data as hierarchical models like generalised linear mixed effect models (glmm).
For example, you might have a multidimensional dataset where your observations have a nested structure (e.g. they are part of the same phylogeny).
You can then analyse this data using a glmm with something like `my_data ~ observations + phylogeny + redisduals`.
For more info on these models [start here](https://en.wikipedia.org/wiki/Generalized_linear_mixed_model).
For more details on running these models, I suggest using the `MCMCglmm` package (@MCMCglmm) from @hadfield2010 (but see also @mulTree).

@@

```{r}
par(mfrow = c(2,2))
## The traitspace
covar.plot(my_covar, col = c("orange", "darkgreen", "blue"), main = "Trait space")
## The traitspace's variance-covariance mean ellipses
covar.plot(my_covar, col = c("orange", "darkgreen", "blue", "grey"), main = "Mean VCV ellipses",
           points = FALSE, ellipses = mean) 
## The traitspace's variance-covariance mean ellipses
covar.plot(my_covar, col = c("orange", "darkgreen", "blue", "grey"), main = "Mean major axes",
           points = FALSE, major.axes = mean)
## A bit of everything
covar.plot(my_covar, col = c("orange", "darkgreen", "blue", "grey"), main = "Ten random VCV matrices",
           points = TRUE, major.axes = TRUE, points.cex = 1/3, n = 10, ellipses = TRUE, legend = TRUE)
```

>>>>

@@

```{r, fig.width=12, fig.height=6}
## Making the different subsets
cust_subsets <- custom.subsets(BeckLee_mat99,
                               crown.stem(BeckLee_tree,
                                          inc.nodes = TRUE))
time_subsets <- chrono.subsets(BeckLee_mat99,
                               tree = BeckLee_tree,
                               method = "discrete",
                               time = 5)

@@

```{r, fig.height = 12, fig.width = 6}
par(mfrow = c(2,1))
## Default plotting
plot(disparity_time_slices, main = "Disparity through time")
## Plotting with more arguments
plot(disparity_time_slices, type = "preview",
     main = "Two first dimensions of the trait space")
```

@@

```{r, fig.height = 12, fig.width = 12}
## Loading some demo data:
## An ordinated matrix with node and tip labels
data(BeckLee_mat99)
## The corresponding tree with tip and node labels
data(BeckLee_tree)
## A list of tips ages for the fossil data
data(BeckLee_ages)

@@

```{r, fig.width=8, fig.height=4}
## Graphical parameters
op <- par(bty = "n", mfrow = c(1, 2))

@@

```{r, print = FALSE, fig.width = 4, fig.height = 12}
par(mfrow = c(3,1))
slices <- c(7.9, 3.95, 0)
fun.plot <- function(tree) {
  plot(tree)
  nodelabels(tree$node.label, cex = 0.8)
  axisPhylo()
  abline(v = tree$root.time - slices)
}
silent <- lapply(trees, fun.plot)
```

@@

Variance-covariance matrices are sometimes a useful way to summarise multidimensional data.
In fact, you can express the variation in your multidimensional dataset directly in terms of how your trait covary rather than simply the positions of your elements in the trait space.
Furthermore, variance-covariance matrices can be estimated from multidimensional in sometimes more useful ways that simply looking at the the data in your trait space.
This can be done by describing your data as hierarchical models like generalised linear mixed effect models (glmm).
For example, you might have a multidimensional dataset where your observations have a nested structure (e.g. they are part of the same phylogeny).
You can then analyse this data using a glmm with something like `my_data ~ observations + phylogeny + redisduals`.
For more info on these models [start here](https://en.wikipedia.org/wiki/Generalized_linear_mixed_model).
For more details on running these models, I suggest using the `MCMCglmm` package (@MCMCglmm) from @hadfield2010 (but see also @mulTree).
For an example use of this code, see @guillerme2023innovation.

@@

```{r, fig.height = 12, fig.width = 12}
par(mfrow = c(2,2))
## The traitspace
covar.plot(my_covar, col = c("orange", "darkgreen", "blue"), main = "Trait space")
## The traitspace's variance-covariance mean ellipses
covar.plot(my_covar, col = c("orange", "darkgreen", "blue", "grey"), main = "Mean VCV ellipses",
           points = FALSE, ellipses = mean) 
## The traitspace's variance-covariance mean ellipses
covar.plot(my_covar, col = c("orange", "darkgreen", "blue", "grey"), main = "Mean major axes",
           points = FALSE, major.axes = mean)
## A bit of everything
covar.plot(my_covar, col = c("orange", "darkgreen", "blue", "grey"), main = "Ten random VCV matrices",
           points = TRUE, major.axes = TRUE, points.cex = 1/3, n = 10, ellipses = TRUE, legend = TRUE)
```

### 2024-11-12 inst/gitbook/05_other-functionalities.Rmd
book up

```{r}
set.seed(42)
## Generating simple coalescent tree
tree <- rcoal(5)

@@

```{r}
set.seed(42)
## Generating a tree
tree <- rtree(20)

>>>>

```{r, fig.height = 12, fig.width = 4}
set.seed(42)
## Generating simple coalescent tree
tree <- rcoal(5)

@@

```{r, fig.height = 12, fig.width = 4}
set.seed(42)
## Generating a tree
tree <- rtree(20)
